---
title: "Untitled"
author: "James Wolpe"
date: "3/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

All work presented is my own. I have not communicated with or worked with anyone else on this exam.
James Wolpe

##Introduction 
The goal of this tutorial is to show a very useful tool for data visualization: mapping. The data I used is of my home state New Hampshire 

The maps package provides great maps

```{r}
#install.packages("maps")
library(maps)
```
However I want to make a map of a less common area so I will be using something called a shapefile. A few links in the course notes are useful for finding pre-made shapefiles, I selected one on New Hampshire from the census website. 

To use shapefiles we will have to use a package called "sf", which stands for "simple features" (not shapefiles).
```{r}
##install.packages("sf")
library(sf)
library(tidyverse)
```

We will now read in the shapefile and use the st_transform command to make the data workable.
```{r}
shp <- read_sf("cb_2018_33_bg_500k/cb_2018_33_bg_500k.shp") %>%
  st_transform(st_crs("+proj=longlat"))
```

Now we can draw a map, to do this we use the geom_sf function.
```{r}
library(tidyverse)
ggplot(data = shp) + 
  geom_sf() +
  theme_void()
```

Our map is now drawn and we can see that each town is outlined (none compare to the beauty of Dublin and Mount Monadnock), however to make this a little more useful, lets try to fill in each county by color.
```{r}
ggplot(data = shp) +
  geom_sf(aes(fill = COUNTYFP)) +
  theme_void()
```

Success! This covers the basics of how to create a shapefile map. Of all the visualizations mapping is perhaps the most beautiful.