---
title: "Quiz 1 in class"
output: html_document
---



1. the easiest catgory to view the overall trend would be group 5 as it has a baseline of zero so they can all be evaluated on the same scale.

2.
```{r}
library(tidyverse)
category <- rep(c("Group 1", "Group 2", "Group 3", "Group 4", "Group 5"), 7)
time <- rep(c(2000, 2001, 2002, 2003, 2004, 2005, 2006), each = 5) 
response <- rpois(length(category), 30)
df <- tibble(category, time, response)
ggplot(data = df, aes(x = time, y = response)) +
  geom_col(aes(fill = category)) +
  facet_wrap(~ response) +
  labs(y = "count")
```

3a. We would use diverging scale as it would have a clear neutral baseline showing no change and would be easy to see positive or negitive change based on the intesity of the two hues.

3b. We would use sequential as the gradient can represnt high or low covid cases clearly, ie high covid cases might be bright red while medium would be a lighter grey-red.

```{r}
ggplot(data = df, aes(x = time, y = response)) +
  geom_col(aes(fill = category)) +
  coord_flip()
```

4. The following plot modifies mapping using the aes function to modify fill and coordinate function coord_flip to stack the bar horizontally.

5a.i

```{r}
ggplot(data = df, aes(x = time, y = response)) +
  geom_point(colour = "blue")
```
You would need to just put color = blue in the geom function so it effects the entire plot.

5b

```{r}
ggplot(data = df, aes(x = time, y = response, colour = category)) +
  geom_point() +
  geom_smooth(se = FALSE, span = 1.9)
```
